---
title: "p8105_hw6_yh3676"
author: "Yuzhe Hu"
date: "2023-12-02"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(viridis)
library(modelr)
library(readr)
library(ggplot2)
set.seed(321)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
	dpi = 300,
  out.width = "90%"
)

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

### Problem 1

In the data cleaning code below we create a `city_state` variable, change `victim_age` to numeric, modifiy victim_race to have categories white and non-white, with white as the reference category, and create a `resolution` variable indicating whether the homicide is solved. Lastly, we filtered out the following cities: Tulsa, AL; Dallas, TX; Phoenix, AZ; and Kansas City, MO; and we retained only the variables `city_state`, `resolution`, `victim_age`, `victim_sex`, and `victim_race`.

```{r q1_data_cleaning, message=FALSE}
homicide_df = 
  read_csv("data/homicide-data.csv", na = c("", "NA", "Unknown")) |> 
  mutate(
    city_state = str_c(city, state, sep = ", "),
    victim_age = as.numeric(victim_age),
    resolution = case_when(
      disposition == "Closed without arrest" ~ 0,
      disposition == "Open/No arrest"        ~ 0,
      disposition == "Closed by arrest"      ~ 1)
  ) |> 
  filter(victim_race %in% c("White", "Black")) |> 
  filter(!(city_state %in% c("Tulsa, AL", "Dallas, TX", "Phoenix, AZ", "Kansas City, MO"))) |> 
  select(city_state, resolution, victim_age, victim_sex, victim_race)
```

Next we fit a logistic regression model using only data from Baltimore, MD. We model `resolved` as the outcome and `victim_age`, `victim_sex`, and `victim_race` as predictors. We save the output as `baltimore_glm` so that we can apply `broom::tidy` to this object and obtain the estimate and confidence interval of the adjusted odds ratio for solving homicides comparing non-white victims to white victims.

```{r q1_glm_baltimore}
baltimore_glm = 
  filter(homicide_df, city_state == "Baltimore, MD") |> 
  glm(resolution ~ victim_age + victim_sex + victim_race, family = binomial(), data = _)

baltimore_glm |> 
  broom::tidy() |> 
  mutate(
    OR = exp(estimate), 
    OR_CI_upper = exp(estimate + 1.96 * std.error),
    OR_CI_lower = exp(estimate - 1.96 * std.error)) |> 
  filter(term == "victim_sexMale") |> 
  select(OR, OR_CI_lower, OR_CI_upper) |>
  knitr::kable(digits = 3)
```

Below, by incorporating `nest()`, `map()`, and `unnest()` into the preceding Baltimore-specific code, we fit a model for each of the cities, and extract the adjusted odds ratio (and CI) for solving homicides comparing non-white victims to white victims. We show the first 5 rows of the resulting dataframe of model results.

```{r q1_glm_all_cities}
model_results = 
  homicide_df |> 
  nest(data = -city_state) |> 
  mutate(
    models = map(data, \(df) glm(resolution ~ victim_age + victim_sex + victim_race, 
                             family = binomial(), data = df)),
    tidy_models = map(models, broom::tidy)) |> 
  select(-models, -data) |> 
  unnest(cols = tidy_models) |> 
  mutate(
    OR = exp(estimate), 
    OR_CI_upper = exp(estimate + 1.96 * std.error),
    OR_CI_lower = exp(estimate - 1.96 * std.error)) |> 
  filter(term == "victim_sexMale") |> 
  select(city_state, OR, OR_CI_lower, OR_CI_upper)

model_results |>
  slice(1:5) |> 
  knitr::kable(digits = 3)
```

Below we generate a plot of the estimated ORs and CIs for each city, ordered by magnitude of the OR from smallest to largest. From this plot we see that most cities have odds ratios that are smaller than 1, suggesting that crimes with male victims have smaller odds of resolution compared to crimes with female victims after adjusting for victim age and race. This disparity is strongest in New yrok. In roughly half of these cities, confidence intervals are narrow and do not contain 1, suggesting a significant difference in resolution rates by sex after adjustment for victim age and race. 

```{r q1_plot}
model_results |> 
  mutate(city_state = fct_reorder(city_state, OR)) |> 
  ggplot(aes(x = city_state, y = OR)) + 
  geom_point() + 
  geom_errorbar(aes(ymin = OR_CI_lower, ymax = OR_CI_upper)) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

### Problem 2
Here we will focus on a simple linear regression with `tmax` as the response with `tmin` and `prcp` as the predictors, and are interested in the distribution of two quantities estimated :
$\hat{r}^2$ and $log (\hat{β}_1 * \hat{β}_2)$. 
Use 5000 bootstrap samples and, for each bootstrap sample, produce estimates of these two quantities.
```{r data, message = FALSE}
#import data
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2022-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  select(name, id, everything())
```

Generate 5000 bootstrap samples
```{r}
# bootstrap function
boot_sample = function(df) {
  
  sample_frac(df, replace = TRUE) 
  
}
# draw the 5000 samples
boot_straps =
  tibble(strap_number = 1:5000) |>
  mutate(
    strap_sample = map(strap_number, \(i) boot_sample(weather_df))
  )
```

Fit the linear regression model
```{r}
boot_results =
  boot_straps |>
  mutate(
    # map each sample to lm() to get linear models
    models = map(strap_sample, \(df) lm(tmax ~ tmin + prcp, data = df)), 
    glance = map(models, broom::glance),
    results = map(models, broom::tidy),
    r_squared = map(glance, \(df) pull(df, r.squared)),
    log_b1b2 = map(results, \(df) log(abs(df$estimate[2] * df$estimate[3])))
  ) |>
  select(strap_number, r_squared, log_b1b2) |>
  mutate(
    r_squared = as.numeric(r_squared),
    log_b1b2 = as.numeric(log_b1b2)
  )
```

Plot the estimates
```{r}
#plot the distribution of r-squared
boot_results |>
  ggplot(aes(x = r_squared)) +
  geom_density() +
  labs(title = "Distribution of R-squared")
```

* The $\hat{r}^2$ distribution is centered around 0.92, and most of the $\hat{r}^2$ values exceed 0.85, which means the current model with `tmin` and `prcp` as the predictors explains a large proportion of the variance in the response variable `tmax`.

```{r}
#plot the distribution of log(beta1_hat * beta2_hat)
boot_results |>
  ggplot(aes(x = log_b1b2)) +
  geom_density() +
  labs(title = "Distribution of log(beta1_hat * beta2_hat)")
```

* The $log (\hat{β}_1 * \hat{β}_2)$ distribution is centered around -5 but left-skewed.This might because many outliers are included in the bootstrap sample.

Provide 95% confidence interval for $\hat{r}^2$ and $log (\hat{β}_1 * \hat{β}_2)$ using 5000 bootstrap samples

```{r}
# CI for R_square
boot_results |>
  summarize(
    R_squared_CI_lower = quantile(r_squared, 0.025),
    R_squared_CI_upper = quantile(r_squared, 0.975)) |>
  knitr::kable(digits = 3)
```

```{r}
# CI for log(beta1_hat * beta2_hat)
boot_results |>
  summarize(
    log_b1b2_CI_lower = quantile(log_b1b2, 0.025),
    log_b1b2_CI_upper = quantile(log_b1b2, 0.975)) |>
  knitr::kable(digits = 3)
```

### Problem 3

Load and clean the data for regression analysis (i.e. convert numeric to factor where appropriate, check for missing data, etc.)

```{r, message=FALSE}
birthwt_df = 
  read_csv("data/birthweight.csv") |>
  janitor::clean_names() |>
#convert some variables' type into factor
  mutate(babysex = as.factor(babysex),
         frace = as.factor(frace),
         malform = as.factor(malform),
         mrace = as.factor(mrace)) |>
  select(bwt,everything())
```

```{r}
#check for missing data
is.na(birthwt_df) |> 
colSums()
```

First, establish a linear regression model with birthweight as the response variable, and use all the other variables as covariates. 

```{r}
full_fit = lm(bwt ~., data = birthwt_df) 
broom::tidy(full_fit) |> 
  knitr::kable()
```

Then, select predictors through the backward stepwise method.

```{r}
stepwise_model = MASS::stepAIC(full_fit, trace = FALSE, direction = "backward")
broom::tidy(stepwise_model) |> 
  knitr::kable()
```

* The selected predictors are: `babysex`, `bhead`, `blength`, `delwt`, `fincome`, `gaweeks`, `mheight`, `mrace`, `parity`, `ppwt`, `smoken`.

Plot of model residuals against fitted value

